test

### 2023-10-16 ms/ms.Rmd
added model_var_table

```{r}
source("r/header.R")
fit = readr::read_rds("objects/fit.rds")
```

@@

We modeled the coordination advantage over a range of leaf thicknesses, stomatal densities, photosynthetic capacities, and light environments to understand when offsetting stomatal position on each surface might deliver a significant photosynthetic advantage (Table X).

>>>>

```{r}
source("r/header.R")
fit = readr::read_rds("objects/fit.rds")
model_var_table = readr::read_rds("objects/model_var_table.rds")
```

@@

We modeled the coordination advantage over a range of leaf thicknesses, stomatal densities, photosynthetic capacities, and light environments to understand when offsetting stomatal position on each surface might deliver a significant photosynthetic advantage (Table \ref{tab:model_var}).

```{r, model_var}
knitr::kable(model_var_table)
```

### 2023-10-16 ms/ms.tex
added model_var_table

\emph{Arabidopsis thaliana} plants were grown in three different light
environments: low light (50 PAR), medium light (100 PAR), and high light
(200 PAR)\ldots{} Once leaves were mature, we captured images of the
abaxial and adaxial leaf surfaces using xxx bifocal microscope. This
microscope allows the capture of two focal lengths that are spatially
correlated such that pixel (1,1) of the abaxial surface image is
directly above pixel (1,1) of the adaxial surface image, and so
on\ldots{} Images were 512 by 512 pixels and covered X leaf surface
area. We captured 132 images in total, making 66 abaxial-adaxial image
pairs. The position of all stomata were recorded using ImageJ
(\protect\hyperlink{ref-schneider_nih_2012}{Schneider, Rasband, and
Eliceiri 2012}).

@@

We tested whether stomata are non-randomly distributed by comparing the
observed stomatal patterning to a random uniform pattern. For each leaf
surface image with \(n\) stomata we generated \(10^3\) synthetic
surfaces with \(n\) stomata uniformly randomly distributed on the
surface. For each sample image, we compared the observed Nearest
Neighbor Index (\(\mathrm{NNI}\)) to the null distribution of
\(\mathrm{NNI}\) values calculated from the synthetic data set. The
observed stomatal distribution is dispersed relative to a uniform random
distribution if the observed \(NNI\) is greater than 95\% of the
synthetic \(\mathrm{NNI}\) values (one-tailed test).

\(\mathrm{NNI}\) is the ratio of observed mean distance
(\(\overline{D}_O\)) to the expected mean distance (\(\overline{D}_E\))
where \(\overline{D}_E\) is:

\begin{equation}\label{eq:emd}
  \overline{D}_E = \frac{0.5}{\sqrt{A_\text{leaf} / n_\text{stomata}}}
\end{equation}

\noindent \(A_\text{leaf}\) is leaf area visible in the sampled field
and \(n_\text{stomata}\) the number of stomata. \(\overline{D}_E\) is
the theoretical average distance to the nearest neighbor of each stomate
if stomata were uniformly randomly distributed
(\protect\hyperlink{ref-clark_distance_1954}{Clark and Evans 1954}). And
\(\overline{D}_O\) calculated for each synthetic data set is:

\begin{equation}\label{eq:omd}
  \overline{D}_O = \frac{\sum_{i=1}^{n_\text{stomata}}d_i}{n_\text{stomata}}
\end{equation}

\(d_i\) is the distance between \(\text{stomate}_i\) and its nearest
neighbor. We calculated \(\mathrm{NNI}\) using the R package
\textbf{spatialEco} version 2.0.1
(\protect\hyperlink{ref-evans_spatialeco_2023}{J. S. Evans and Murphy
2023}).

@@

\noindent \(\mathrm{NNI}\) is calculated for each sample image as
described above; \(\text{median}(\mathrm{NNI_{random}})\) and
\(\text{median}(\mathrm{NNI_{uniform}})\) are calculated from the dry og
synthetic data specific to each sample image as described above. We
tested whether light treatment affects \(\mathrm{DI}\) and \(D_S\) using
analysis of variance (ANOVA).

Finally, we examined the relationship between stomatal zone area and
stomatal length using a Bayesian generalized non-linear multilevel model
with the R package \textbf{brms} version 2.20.1
(\protect\hyperlink{ref-burkner_brms_2017}{Bürkner 2017},
\protect\hyperlink{ref-burkner_advanced_2018}{2018}). Stomatal zone area
was calculated using Voronoi tessellation (e.g.~Fig. \ref{fig:example}).
The stomatal zone area, \(S_\text{area}\), is the region of the leaf
surface whose distance to stomate, \(S\), is less than the distance to
any other stomate, \(S\). Stomatal length was measured in
\texttt{ImageJ} (\protect\hyperlink{ref-schneider_nih_2012}{Schneider,
Rasband, and Eliceiri 2012}).

@@

To test whether the position of ab- and adaxial stomata are coordinated
we compared the observed distribution to a null distribution where the
positions on each surface are random. For each pair of surfaces
(observed or synthetic) we calculated the distance squared between each
to the nearest stomatal centroid with the R package \textbf{raster}
version 3.6.23. Then we calculated the cell-wise Pearson correlation
coefficient. If stomatal positions on each surface are coordinated to
minimize the distance between mesophyll and the nearest stomate, then we
expect a negative correlation. A cell that is far from a stomate on one
surface should be near a stomate on the other surface (Fig.
\ref{fig:amphi}). We generated a null distribution of the correlation
coefficient by simulating \(10^3\) synthetic data sets for each observed
pair. For each synthetic data set, we simulated stomatal position using
a random uniform distribution, as described above, matching the number
of stomata on abaxial and adaxial leaf surfaces. Stomatal positions on
each surface are coordinated if the correlation coefficient is greater
than 95\% of the synthetic correlation values (one-tailed test).

@@

We modeled photosynthesis CO\(_2\) assimilation rate using a
spatially-explicit two-dimensional reaction diffusion model using a
porous medium approximation (Parkhurst 1994). Consider a two-dimensional
leaf where stomata occur on each surface in a regular sequence with
interstomatal distance \(U\). The main outcome we assessed is the
advantage of offsetting the position of stomata on each surface compared
to have stomata on the same \(x\) position on each surface. With these
assumptions, by symmetry, we only need to model two stomata, one abaxial
and one adaxial, from \(x = 0\) to \(x = U/2\) and from the adaxial
surface at \(y = 0\) to the abaxial surface at \(y = L\), the leaf
thickness. We arbitrarily set the adaxial stomate at \(x = 0\) and
toggled the abaxial stomata position between \(x = U/2\) (offset) or
\(x = 0\) (below adaxial stomate). The advantage of offsetting stomatal
position on each surface is the photosynthetic rate of the leaf with
offset stomata compared to that with stomata aligned in the same \(x\)
position:

@@

We modelled the coordination advantage over a range of leaf thicknesses,
stomatal densities, photosynthetic capacities, and light environments to
understand when offsetting stomatal position on each surface might
deliver a significant photosynthetic advantage.

@@

{[}NOTE: we should probably adjust light attenuation to be proportional
to the chlorophyll concentration which is one of the parameters in the
biochemical models.{]}

@@

FYI - something is wrong with the model. I think the units are off
somewhere because the result is very sensitive to grid size and the
values often aren't reasonable. I'll need to troubleshoot more, but
maybe something will be obvious to you.

We used the \texttt{steady.2D()} function in the R package
\textbf{rootSolve}. The function needs a function to calculate the time
differential for node \(ij\), \(\frac{dC_{m,ij}}{dt}\). The R function
below uses the light propagation, biochemical, and diffusional model to
calculate a matrix of \(\frac{dC_{m,ij}}{dt}\) values:

>>>>

\emph{Arabidopsis thaliana} plants were grown in three different light
environments: low light (PAR = 50
\(\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}\)), medium light (100
\(\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}\)), and high light (200
\(\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}\)). PAR stands for
photosynthetically active radiation. Once leaves were mature, we
captured images of the abaxial and adaxial leaf surfaces using XXX. We
captured 132 images in total, making 66 abaxial-adaxial image pairs. We
measured stomatal position and size using ImageJ
(\protect\hyperlink{ref-schneider_nih_2012}{Schneider, Rasband, and
Eliceiri 2012}).

@@

We tested whether stomata are non-randomly distributed by comparing the
observed stomatal patterning to a random uniform pattern. For each leaf
surface image with \(n\) stomata we generated \(10^3\) synthetic
surfaces with \(n\) stomata uniformly randomly distributed on the
surface. For each sample image, we compared the observed Nearest
Neighbor Index (\(\mathrm{NNI}\)) to the null distribution of
\(\mathrm{NNI}\) values calculated from the synthetic data set.
\(\mathrm{NNI}\) is the ratio of observed mean distance
(\(\overline{D}_O\)) to the expected mean distance (\(\overline{D}_E\))
where \(\overline{D}_E\) is:

\begin{equation}\label{eq:emd}
  \overline{D}_E = \frac{0.5}{\sqrt{A_\text{leaf} / n_\text{stomata}}}.
\end{equation}

\noindent \(A_\text{leaf}\) is leaf area visible in the sampled field
and \(n_\text{stomata}\) the number of stomata. \(\overline{D}_E\) is
the theoretical average distance to the nearest neighbor of each stomate
if stomata were uniformly randomly distributed
(\protect\hyperlink{ref-clark_distance_1954}{Clark and Evans 1954}).
\(\overline{D}_O\) calculated for each synthetic data set is:

\begin{equation}\label{eq:omd}
  \overline{D}_O = \frac{\sum_{i=1}^{n_\text{stomata}}d_i}{n_\text{stomata}},
\end{equation}

\noindent where \(d_i\) is the distance between \(\text{stomate}_i\) and
its nearest neighbor. We calculated \(\mathrm{NNI}\) using the \emph{R}
package \textbf{spatialEco} version 2.0.1
(\protect\hyperlink{ref-evans_spatialeco_2023}{J. S. Evans and Murphy
2023}). The observed stomatal distribution is dispersed relative to a
uniform random distribution if the observed \(NNI\) is greater than 95\%
of the synthetic \(\mathrm{NNI}\) values (one-tailed test).

@@

\noindent \(\mathrm{NNI}\) is calculated for each sample image as
described above; \(\text{median}(\mathrm{NNI_{random}})\) and
\(\text{median}(\mathrm{NNI_{uniform}})\) are calculated from the
synthetic data specific to each sample image as described above. We
tested whether light treatment affects \(\mathrm{DI}\) and stomatla
density (\(D_S\)) using analysis of variance (ANOVA).

Finally, we examined the relationship between stomatal zone area and
stomatal length using a Bayesian generalized non-linear multilevel model
fit with the \emph{R} package \textbf{brms} version 2.15.0
(\protect\hyperlink{ref-burkner_brms_2017}{Bürkner 2017},
\protect\hyperlink{ref-burkner_advanced_2018}{2018}) and \emph{Stan}
version 2.27.0
(\protect\hyperlink{ref-stan_development_team_stan_2023}{\textbf{stan\_development\_team\_stan\_2023?}}).
Stomatal zone area was calculated using Voronoi tessellation (e.g.~Fig.
\ref{fig:example}). The stomatal zone area, \(S_\text{area}\), is the
region of the leaf surface whose distance to stomate, \(S\), is less
than the distance to any other stomate, \(S\). Stomatal length was
measured in ImageJ
(\protect\hyperlink{ref-schneider_nih_2012}{Schneider, Rasband, and
Eliceiri 2012}).

@@

To test whether the position of ab- and adaxial stomata are coordinated
we compared the observed distribution to a null distribution where the
positions on each surface are random. For each pair of surfaces
(observed or synthetic) we calculated the distance squared between each
to the nearest stomatal centroid with the \emph{R} package
\textbf{raster} version 3.6.23. Then we calculated the cell-wise Pearson
correlation coefficient. If stomatal positions on each surface are
coordinated to minimize the distance between mesophyll and the nearest
stomate, then we expect a negative correlation. A cell that is far from
a stomate on one surface should be near a stomate on the other surface
(Fig. \ref{fig:amphi}). We generated a null distribution of the
correlation coefficient by simulating \(10^3\) synthetic data sets for
each observed pair. For each synthetic data set, we simulated stomatal
position using a random uniform distribution, as described above,
matching the number of stomata on abaxial and adaxial leaf surfaces.
Stomatal positions on each surface are coordinated if the correlation
coefficient is greater than 95\% of the synthetic correlation values
(one-tailed test).

@@

We modeled photosynthesis CO\(_2\) assimilation rate using a
spatially-explicit two-dimensional reaction diffusion model using a
porous medium approximation
(\protect\hyperlink{ref-parkhurst_intercellular_1994}{\textbf{parkhurst\_intercellular\_1994?}})
using the finite element method (FEM) following
(\protect\hyperlink{ref-earles_excess_2017}{\textbf{earles\_excess\_2017?}}).
Consider a two-dimensional leaf where stomata occur on each surface in a
regular sequence with interstomatal distance \(U\). The main outcome we
assessed is the advantage of offsetting the position of stomata on each
surface compared to have stomata on the same \(x\) position on each
surface. With these assumptions, by symmetry, we only need to model two
stomata, one abaxial and one adaxial, from \(x = 0\) to \(x = U/2\) and
from the adaxial surface at \(y = 0\) to the abaxial surface at
\(y = L\), the leaf thickness. We arbitrarily set the adaxial stomate at
\(x = 0\) and toggled the abaxial stomata position between \(x = U/2\)
(offset) or \(x = 0\) (below adaxial stomate). The `coordination
advantage' of offsett stomatal position on each surface is the
photosynthetic rate of the leaf with offset stomata compared to that
with stomata aligned in the same \(x\) position:

@@

We modeled the coordination advantage over a range of leaf thicknesses,
stomatal densities, photosynthetic capacities, and light environments to
understand when offsetting stomatal position on each surface might
deliver a significant photosynthetic advantage (Table
\ref{tab:model_var}).

\begin{longtable}[]{@{}lll@{}}
\toprule\noalign{}
Variable & Parameter range & Units \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\(I_0\) & \(50-1000\) & \(\mu\)mol m\(^{-2}\) s\(^{-1}\) \\
\(\varphi_\text{pal}\) & \(0.1-0.3\) & m\(^3\) airspace m\(^{-3}\)
leaf \\
\(T_\text{leaf}\) & \(101-501\) & \(\mu\)m \\
\(U\) & \(17-169\) & \(\mu\)m \\
\end{longtable}

WORKING HERE - make table of parameter combinations!

@@

@@

We used the \texttt{steady.2D()} function in the \emph{R} package
\textbf{rootSolve} version 1.8.2.4
(\protect\hyperlink{ref-soetaert_practical_2009}{\textbf{soetaert\_practical\_2009?}}).

### 2023-10-16 ms/test.Rmd
added model_var_table

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r dispersion, fig.cap = "Stomata are more dispersed than expected under the null model of uniform random position (dispersion index = 0) but far from a distribution that maximizes distance between stomata (dispersion index = 1)", echo = TRUE}

knitr::include_graphics("figures/dispersion.pdf")

```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE) 
knitr::opts_chunk$set(fig.pos = 'p') # Places figures on pages separate from text
knitr::opts_chunk$set(out.width = '100%', dpi = 300) # Figure resolution and size
knitr::opts_chunk$set(fig.env = "figure") # Latex figure environment
knitr::opts_knit$set(root.dir = "..")
```

```{r}
source("r/header.R")
fit = readr::read_rds("objects/fit.rds")
model_var_table = readr::read_rds("objects/model_var_table.rds")
```

```{r, model_var}
knitr::kable(model_var_table)
```
